---
title: "Combined Samples - 07 Cell-type Abundance"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path(
      "..", "..", "results", "comb", "comb_07abundance.html"
    ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 25px; text-indent: 0;
}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-snRNAseq](https://github.com/eturkes/tau-mutant-snRNAseq).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-snRNAseq.
#    Copyright (C) 2024-2025  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
packages <- c("Seurat", "ggplot2", "scales", "pals")
invisible(
  suppressPackageStartupMessages(
    lapply(packages, FUN = library, character.only = TRUE)
  )
)
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 96)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "utils.R"))
# -----------------

# Useful variables.
# -----------------
step <- 2

cache_dir <- file.path("..", "..", "cache", "comb", paste0("step", "0", step))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
# -----------------

# Read data.
# ----------
seurat_orig <- readRDS(file.path(cache_dir, "annotated_seurat.rds"))
seurat_orig <- seurat_orig[
  , seurat_orig$genotype == "MAPTKI" | seurat_orig$genotype == "S305N"
]
seurat_orig$genotype <- factor(seurat_orig$genotype)

gene_anno <- data.frame(
  symbol = unlist(seurat_orig@misc), ensembl = rownames(seurat_orig[["RNA"]])
)
# ----------
```

```{r}
```{r}
df <- as.data.frame(table(seurat$ctype, seurat$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = percent, breaks = pretty_breaks(5)) +
  ylab("Percent of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$ctype),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_y_continuous(breaks = pretty_breaks(5)) +
  ylab("Number of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$ctype),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
df <- as.data.frame(table(seurat$ctype, seurat$genotype_batch))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = percent, breaks = pretty_breaks(5)) +
  ylab("Percent of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$ctype),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_y_continuous(breaks = pretty_breaks(5)) +
  ylab("Number of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$ctype),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )

df <- as.data.frame(table(seurat$broad_annotations, seurat$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = percent, breaks = pretty_breaks(5)) +
  ylab("Percent of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$broad_annotations),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_y_continuous(breaks = pretty_breaks(5)) +
  ylab("Number of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$broad_annotations),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
df <- as.data.frame(table(seurat$broad_annotations, seurat$genotype_batch))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = percent, breaks = pretty_breaks(5)) +
  ylab("Percent of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$broad_annotations),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_y_continuous(breaks = pretty_breaks(5)) +
  ylab("Number of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$broad_annotations),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )

df <- as.data.frame(table(seurat$level1_subclusters, seurat$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = percent, breaks = pretty_breaks(5)) +
  ylab("Percent of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$level1_subclusters),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_y_continuous(breaks = pretty_breaks(5)) +
  ylab("Number of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$level1_subclusters),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
df <- as.data.frame(table(seurat$level1_subclusters, seurat$genotype_batch))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = percent, breaks = pretty_breaks(5)) +
  ylab("Percent of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$level1_subclusters),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_y_continuous(breaks = pretty_breaks(5)) +
  ylab("Number of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$level1_subclusters),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )

df <- as.data.frame(table(seurat$level2_subclusters, seurat$genotype))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = percent, breaks = pretty_breaks(5)) +
  ylab("Percent of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$level2_subclusters),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_y_continuous(breaks = pretty_breaks(5)) +
  ylab("Number of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$level2_subclusters),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
df <- as.data.frame(table(seurat$level2_subclusters, seurat$genotype_batch))
colnames(df) <- c("Celltype", "Genotype", "Proportion")
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "fill", stat = "identity") +
  scale_y_continuous(labels = percent, breaks = pretty_breaks(5)) +
  ylab("Percent of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$ctype),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
ggplot(df, aes(Genotype, Proportion, fill = Celltype)) +
  geom_bar(position = "stack", stat = "identity") +
  scale_y_continuous(breaks = pretty_breaks(5)) +
  ylab("Number of Cells") +
  coord_flip() +
  theme_classic() +
  scale_fill_manual(
    breaks = levels(seurat$ctype),
    values = as.vector(cols25(length(unique(df$Celltype))))
  )
```
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
