---
title: "Combined Samples - 05 CellChat"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path(
      "..", "..", "results", "comb", "comb_05cellchat.html"
    ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 25px; text-indent: 0;
}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-snRNAseq](https://github.com/eturkes/tau-mutant-snRNAseq).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-snRNAseq.
#    Copyright (C) 2024-2025  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
packages <- c(
  "Seurat", "parallelly", "ggplot2", "dplyr", "ggrepel", "viridis", "DT",
  "CellChat"
)
invisible(
  suppressPackageStartupMessages(
    lapply(packages, FUN = library, character.only = TRUE)
  )
)
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 150)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "utils.R"))
# -----------------

# Useful variables.
# -----------------
step <- 2

data_dir <- file.path("..", "..", "data")
cache_dir <- file.path("..", "..", "cache", "comb", paste0("step", "0", step))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

protocol <- c("mouse", "droplet", "single-nuc", "umis")
vars_to_regress <- NULL
parallel_override <- NULL
# -----------------

# Metadata to plot after dimensionality reduction and clustering.
# Values in list can include "no_legend and/or "no_label" to exclude those.
# -------------------------------------------------------------------------
metadata_to_plot <- vector("list", length = 8)
names(metadata_to_plot) <- c(
  "subcluster_annotations", "allen_labels", "seurat_clusters", "genotype",
  "batch", "genotype_batch", "sex", "Phase"
)
metadata_to_plot[[4]] <- "no_label"
metadata_to_plot[[5]] <- "no_label"
metadata_to_plot[[6]] <- "no_label"
metadata_to_plot[[7]] <- "no_label"
metadata_to_plot[[8]] <- "no_label"
# -------------------------------------------------------------------------
```

# Microglia

```{r}
ctype <- "Microglia"
sub_name <- "micro"

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 800000, dims = 1:24
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Microglia 01",
  "1" = "Microglia 03", "2" = "Microglia 04",
  "3" = "Microglia 02"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

gene_anno <- data.frame(
  symbol = unlist(seurat@misc), ensembl = rownames(seurat[["RNA"]])
)
```

## MAPT KI

```{r, fig.height = 5, fig.width = 5}
seurat_object <- seurat
seurat_object@active.ident <- seurat_object$genotype
seurat_object <- subset(seurat_object, idents = "MAPTKI")
seurat_object$genotype_ctype <- factor(paste(seurat_object$genotype, sub(" ", "", seurat_object$subcluster_annotations), sep = "_"))

data.input <- seurat_object[["SCT"]]$data # normalized data matrix

gene_anno_sub <- gene_anno[gene_anno$ensembl %in% rownames(data.input), ]
rownames(data.input) <- gene_anno_sub$symbol

labels <- seurat_object$genotype_ctype
samples <- seurat_object$genotype
meta <- data.frame(labels = labels, row.names = names(labels), samples = samples) # create a dataframe of the cell labels

cellchat <- createCellChat(object = data.input, meta = meta, group.by = "labels")

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
CellChatDB.use <- subsetDB(CellChatDB)
cellchat@DB <- CellChatDB.use
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multisession", workers = 8) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
cellchat <- smoothData(cellchat, adj = PPI.mouse)
cellchat <- computeCommunProb(cellchat, type = "triMean")
cellchat <- filterCommunication(cellchat, min.cells = 10)
cellchat <- computeCommunProbPathway(cellchat)
cellchat <- aggregateNet(cellchat)

# groupSize <- as.numeric(table(cellchat@idents))
# par(mfrow = c(2,3))
# netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
# netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= T, title.name = "Interaction weights/strength")

for (pathways.show in cellchat@netP$pathways) {
# par(mfrow=c(1,1))
  pairLR <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
  LR.show <- pairLR[1,] # show one ligand-receptor pair
netVisual_individual(cellchat, signaling = pathways.show, big.gap = 0, small.gap = 0, pairLR.use = LR.show)
}
```

## P301S+3

```{r, fig.height = 5, fig.width = 5}
seurat_object <- seurat
seurat_object@active.ident <- seurat_object$genotype
seurat_object <- subset(seurat_object, idents = "P301S")
seurat_object$genotype_ctype <- factor(paste(seurat_object$genotype, sub(" ", "", seurat_object$subcluster_annotations), sep = "_"))

data.input <- seurat_object[["SCT"]]$data # normalized data matrix

gene_anno_sub <- gene_anno[gene_anno$ensembl %in% rownames(data.input), ]
rownames(data.input) <- gene_anno_sub$symbol

labels <- seurat_object$genotype_ctype
samples <- seurat_object$genotype
meta <- data.frame(labels = labels, row.names = names(labels), samples = samples) # create a dataframe of the cell labels

cellchat <- createCellChat(object = data.input, meta = meta, group.by = "labels")

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
CellChatDB.use <- subsetDB(CellChatDB)
cellchat@DB <- CellChatDB.use
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multisession", workers = 8) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
cellchat <- smoothData(cellchat, adj = PPI.mouse)
cellchat <- computeCommunProb(cellchat, type = "triMean")
cellchat <- filterCommunication(cellchat, min.cells = 10)
cellchat <- computeCommunProbPathway(cellchat)
cellchat <- aggregateNet(cellchat)

# groupSize <- as.numeric(table(cellchat@idents))
# par(mfrow = c(2,3))
# netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
# netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= T, title.name = "Interaction weights/strength")

for (pathways.show in cellchat@netP$pathways) {
# par(mfrow=c(1,1))
  pairLR <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
  LR.show <- pairLR[1,] # show one ligand-receptor pair
netVisual_individual(cellchat, signaling = pathways.show, big.gap = 0, small.gap = 0, pairLR.use = LR.show)
}
```

## NLGF MAPT KI

```{r, fig.height = 5, fig.width = 5}
seurat_object <- seurat
seurat_object@active.ident <- seurat_object$genotype
seurat_object <- subset(seurat_object, idents = "NLGF_MAPTKI")
seurat_object$genotype_ctype <- factor(paste(seurat_object$genotype, sub(" ", "", seurat_object$subcluster_annotations), sep = "_"))

data.input <- seurat_object[["SCT"]]$data # normalized data matrix

gene_anno_sub <- gene_anno[gene_anno$ensembl %in% rownames(data.input), ]
rownames(data.input) <- gene_anno_sub$symbol

labels <- seurat_object$genotype_ctype
samples <- seurat_object$genotype
meta <- data.frame(labels = labels, row.names = names(labels), samples = samples) # create a dataframe of the cell labels

cellchat <- createCellChat(object = data.input, meta = meta, group.by = "labels")

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
CellChatDB.use <- subsetDB(CellChatDB)
cellchat@DB <- CellChatDB.use
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multisession", workers = 8) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
cellchat <- smoothData(cellchat, adj = PPI.mouse)
cellchat <- computeCommunProb(cellchat, type = "triMean")
cellchat <- filterCommunication(cellchat, min.cells = 10)
cellchat <- computeCommunProbPathway(cellchat)
cellchat <- aggregateNet(cellchat)

# groupSize <- as.numeric(table(cellchat@idents))
# par(mfrow = c(2,3))
# netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
# netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= T, title.name = "Interaction weights/strength")

for (pathways.show in cellchat@netP$pathways) {
# par(mfrow=c(1,1))
  pairLR <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
  LR.show <- pairLR[1,] # show one ligand-receptor pair
netVisual_individual(cellchat, signaling = pathways.show, big.gap = 0, small.gap = 0, pairLR.use = LR.show)
}
```

## NLGF P301S+3

```{r, fig.height = 5, fig.width = 5}
seurat_object <- seurat
seurat_object@active.ident <- seurat_object$genotype
seurat_object <- subset(seurat_object, idents = "NLGF_P301S")
seurat_object$genotype_ctype <- factor(paste(seurat_object$genotype, sub(" ", "", seurat_object$subcluster_annotations), sep = "_"))

data.input <- seurat_object[["SCT"]]$data # normalized data matrix

gene_anno_sub <- gene_anno[gene_anno$ensembl %in% rownames(data.input), ]
rownames(data.input) <- gene_anno_sub$symbol

labels <- seurat_object$genotype_ctype
samples <- seurat_object$genotype
meta <- data.frame(labels = labels, row.names = names(labels), samples = samples) # create a dataframe of the cell labels

cellchat <- createCellChat(object = data.input, meta = meta, group.by = "labels")

CellChatDB <- CellChatDB.mouse # use CellChatDB.mouse if running on mouse data
CellChatDB.use <- subsetDB(CellChatDB)
cellchat@DB <- CellChatDB.use
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
future::plan("multisession", workers = 8) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
cellchat <- smoothData(cellchat, adj = PPI.mouse)
cellchat <- computeCommunProb(cellchat, type = "triMean")
cellchat <- filterCommunication(cellchat, min.cells = 10)
cellchat <- computeCommunProbPathway(cellchat)
cellchat <- aggregateNet(cellchat)

# groupSize <- as.numeric(table(cellchat@idents))
# par(mfrow = c(2,3))
# netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
# netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= T, title.name = "Interaction weights/strength")

for (pathways.show in cellchat@netP$pathways) {
# par(mfrow=c(1,1))
  pairLR <- extractEnrichedLR(cellchat, signaling = pathways.show, geneLR.return = FALSE)
  LR.show <- pairLR[1,] # show one ligand-receptor pair
netVisual_individual(cellchat, signaling = pathways.show, big.gap = 0, small.gap = 0, pairLR.use = LR.show)
}
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
