---
title: "Combined Samples - 02 Subclustering"
author:
  - name: "Emir Turkes"
date: '`r strftime(Sys.time(), format = "%B %d, %Y")`'
link-citations: true
output:
  html_document:
    code_folding: hide
    number_sections: true
    theme: lumen
    highlight: haddock
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_file = file.path(
      "..", "..", "results", "comb", "comb_02subclustering.html"
    ))})
---

<style type="text/css">
body {font-size: 16px;}
h1.title {font-size: 35px;}
h1 {font-size: 24px;}
h2 {font-size: 22px;}
.toc-content {padding-left: 0px; padding-right: 0px;}
div.tocify {width: 100%;}
.tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 25px; text-indent: 0;
}
.tocify-subheader .tocify-subheader .tocify-item {
  font-size: 0.95em; padding-left: 35px; text-indent: 0;
}
div.main-container {max-width: none; width: 100%;}
</style>

*This file is a part of [tau-mutant-snRNAseq](https://github.com/eturkes/tau-mutant-snRNAseq).*

The table of contents in the top left is clickable and can be used to quickly navigate the document.
To toggle the visibility of code, use the `CODE` toggles at the top right of chunks.
The toggle at the start of the document controls the visibility of all chunks.

```{r}
#    This file is part of tau-mutant-snRNAseq.
#    Copyright (C) 2024  Emir Turkes, Naoto Watamura, UK DRI at UCL
#
#    This program is free software: you can redistribute it and/or modify
#    it under the terms of the GNU General Public License as published by
#    the Free Software Foundation, either version 3 of the License, or
#    (at your option) any later version.
#
#    This program is distributed in the hope that it will be useful,
#    but WITHOUT ANY WARRANTY; without even the implied warranty of
#    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#    GNU General Public License for more details.
#
#    You should have received a copy of the GNU General Public License
#    along with this program.  If not, see <http://www.gnu.org/licenses/>.
#
#    Emir Turkes can be contacted at emir.turkes@eturkes.com

# Load required packages, suppressing startup messages.
# -----------------------------------------------------
library(conflicted)
packages <- c(
  "Seurat", "parallelly", "ggplot2", "dplyr", "ggrepel", "viridis", "DT"
)
invisible(
  suppressPackageStartupMessages(
    lapply(packages, FUN = library, character.only = TRUE)
  )
)
# -----------------------------------------------------

# Define global settings.
# -----------------------
knitr::opts_chunk$set(fig.width = 10, fig.height = 7, dpi = 300)
# -----------------------

# Define functions.
# -----------------
source(file.path("..", "utils.R"))
# -----------------

# Useful variables.
# -----------------
step <- 2

data_dir <- file.path("..", "..", "data")
cache_dir <- file.path("..", "..", "cache", "comb", paste0("step", "0", step))
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}

protocol <- c("mouse", "droplet", "single-nuc", "umis")
vars_to_regress <- NULL
parallel_override <- NULL
# -----------------

# Metadata to plot after dimensionality reduction and clustering.
# Values in list can include "no_legend and/or "no_label" to exclude those.
# -------------------------------------------------------------------------
metadata_to_plot <- vector("list", length = 8)
names(metadata_to_plot) <- c(
  "subcluster_annotations", "allen_labels", "seurat_clusters", "genotype",
  "batch", "genotype_batch", "sex", "Phase"
)
metadata_to_plot[[4]] <- "no_label"
metadata_to_plot[[5]] <- "no_label"
metadata_to_plot[[6]] <- "no_label"
metadata_to_plot[[7]] <- "no_label"
metadata_to_plot[[8]] <- "no_label"
# -------------------------------------------------------------------------
```

# Prep

```{r}
level1_subclusters <- vector("list", length = 6)
level2_subclusters <- vector("list", length = 19)
```

# Neuronal

```{r}
ctype <- "Neuronal"
sub_name <- "neuronal"

seurat <- readRDS(
  file.path(cache_dir, "..", "step01", "processed_merged_seurat.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 2500000, dims = 1:38
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Glutamatergic 01",
  "1" = "Glutamatergic 01", "2" = "Glutamatergic 01",
  "3" = "Glutamatergic 01", "4" = "Glutamatergic 02",
  "5" = "GABAergic 01", "6" = "Glutamatergic 01",
  "7" = "Glutamatergic 03", "8" = "Glutamatergic 01",
  "9" = "GABAergic 01", "10" = "Glutamatergic 01",
  "11" = "Glutamatergic 01", "12" = "GABAergic 01",
  "13" = "Glutamatergic 01", "14" = "Glutamatergic 04",
  "15" = "Glutamatergic 01", "16" = "Glutamatergic 01",
  "17" = "Glutamatergic 05", "18" = "Glutamatergic 01",
  "19" = "Glutamatergic 01"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 38)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level1_subclusters[[1]] <- as.character(seurat$subcluster_annotations)
names(level1_subclusters[[1]]) <- names(seurat$subcluster_annotations)
```

## Glutamatergic 01

```{r}
ctype <- "Glutamatergic 01"
sub_name <- "glut01"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_neuronal.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 357500, dims = 1:47
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Glut DG 01",
  "1" = "Glutamatergic 03", "2" = "Glut L2/3 Cortex 01",
  "3" = "Glut CA1 01", "4" = "Glut L4/5 Cortex 01",
  "5" = "Glutamatergic 05", "6" = "Glutamatergic 01",
  "7" = "Glutamatergic 08", "8" = "Glut CA3 01",
  "9" = "Glut L5/6 Cortex 02", "10" = "Glut L5/6 Cortex 01",
  "11" = "Glutamatergic 04", "12" = "Glut L3 ENT 01",
  "13" = "Glut Subiculum 01", "14" = "Glutamatergic 07",
  "15" = "Glut L5/6 Cortex 05", "16" = "Glutamatergic 11",
  "17" = "Glut L2 ENT 01", "18" = "Glut CA3 02",
  "19" = "Glutamatergic 09", "20" = "Glutamatergic 10",
  "21" = "Glutamatergic 06", "22" = "Glut Car3 01",
  "23" = "Glut L5/6 Cortex 03", "24" = "Glut DG 02",
  "25" = "Glut Cajal-Retzius 01", "26" = "Glut CA3 03",
  "27" = "Glut L5/6 Cortex 04", "28" = "Glut Retrosplenial 01",
  "29" = "Glutamatergic 02", "30" = "Glut Retrohippocampal 01",
  "31" = "Glut L5/6 Cortex 06"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 47)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[1]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[1]]) <- names(seurat$subcluster_annotations)
```

## Glutamatergic 02

```{r}
ctype <- "Glutamatergic 02"
sub_name <- "glut02"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_neuronal.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 60000, dims = 1:27
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Glut L6 Cortex 01",
  "1" = "Glut L6 ENT 01", "2" = "Glut L6 ENT 02", 
  "3" = "Glut Subiculum 02", "4" = "Glut L6 Cortex 04",
  "5" = "Glut L6 Cortex 02", "6" = "Glut L6 Cortex 03", 
  "7" = "Glutamatergic 12", "8" = "Glut L6 PPP 01",
  "9" = "Glutamatergic 13"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 27)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[2]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[2]]) <- names(seurat$subcluster_annotations)
```

## Glutamatergic 03

```{r}
ctype <- "Glutamatergic 03"
sub_name <- "glut03"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_neuronal.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 100000, dims = 1:19
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Glut L2/3 PPP 01",
  "1" = "Glut L2/3 PPP 02", "2" = "Glut L2/3 PPP 05",
  "3" = "Glut L2/3 PPP 04", "4" = "Glut L2/3 PPP 03",
  "5" = "Glut L2/3 PPP 06"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 19)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[3]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[3]]) <- names(seurat$subcluster_annotations)
```

## Glutamatergic 04

```{r}
ctype <- "Glutamatergic 04"
sub_name <- "glut04"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_neuronal.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 100000, dims = 1:16
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Glut Subiculum 03",
  "1" = "Glut L6 Cortex 05"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 16)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[4]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[4]]) <- names(seurat$subcluster_annotations)
```

## Glutamatergic 05

```{r}
ctype <- "Glutamatergic 05"
sub_name <- "glut05"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_neuronal.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 10000, dims = 1:30
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Glut DG 03",
  "1" = "Glut DG 04", "2" = "Glutamatergic 14"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 30)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[5]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[5]]) <- names(seurat$subcluster_annotations)
```

## GABAergic 01

```{r}
ctype <- "GABAergic 01"
sub_name <- "gaba01"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_neuronal.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 200000, dims = 1:27
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "GABA Pvalb 01",
  "1" = "GABA Sst 01", "2" = "GABA Vip 01",
  "3" = "GABAergic 01", "4" = "GABAergic 02",
  "5" = "GABA Lamp5 01", "6" = "GABA Sncg 01",
  "7" = "GABA Lamp5 02", "8" = "GABAergic 03",
  "9" = "GABA Pvalb 02", "10" = "GABAergic 05",
  "11" = "GABAergic 04"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 27)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[6]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[6]]) <- names(seurat$subcluster_annotations)
```

# Microglia

```{r}
ctype <- "Microglia"
sub_name <- "micro"

seurat <- readRDS(
  file.path(cache_dir, "..", "step01", "processed_merged_seurat.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 800000, dims = 1:24
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Microglia 01",
  "1" = "Microglia 03", "2" = "Microglia 04",
  "3" = "Microglia 02"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 24)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level1_subclusters[[2]] <- as.character(seurat$subcluster_annotations)
names(level1_subclusters[[2]]) <- names(seurat$subcluster_annotations)
```

## Microglia 01

```{r}
ctype <- "Microglia 01"
sub_name <- "micro01"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_micro.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 170000, dims = 1:19
)
seurat

file_list <- list.files(file.path(data_dir, "DAM_genes"))
DAM_list <- vector("list", length = length(file_list))
names(DAM_list) <- sapply(file_list, function(x) sub("_.*", "", x))
for (i in seq_along(DAM_list)) {
  DAM_list[[i]] <- read.delim(
    file.path(data_dir, "DAM_genes", file_list[i]), header = FALSE
  )
  gene_anno_sub <- gene_anno[
    tolower(gene_anno$geneid) %in% tolower(DAM_list[[i]]$V1),
  ]
  DAM_list[[i]] <- gene_anno_sub$gene
}

seurat <- AddModuleScore(seurat, features = DAM_list, name = "DAM_genes")
for (i in seq_along(DAM_list)) {
  print(
    FeaturePlot(
      seurat, features = paste0("DAM_genes", i),
      order = TRUE, cols = c("lightgrey", "red")
    ) + ggtitle(names(DAM_list)[i])
  )
}

seurat <- RenameIdents(
  object = seurat, "0" = "Micro Homeostatic 01",
  "1" = "Micro DAM 01", "2" = "Micro DAM 02",
  "3" = "Microglia 01", "4" = "Microglia 02"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 19)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[7]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[7]]) <- names(seurat$subcluster_annotations)
```

## Microglia 02

```{r}
ctype <- "Microglia 02"
sub_name <- "micro02"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_micro.rds")
)
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- RenameIdents(
  object = seurat, "Microglia 02" = "Microglia 03"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[8]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[8]]) <- names(seurat$subcluster_annotations)
```

## Microglia 03

```{r}
ctype <- "Microglia 03"
sub_name <- "micro03"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_micro.rds")
)
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- RenameIdents(
  object = seurat, "Microglia 03" = "Microglia 04"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[9]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[9]]) <- names(seurat$subcluster_annotations)
```

## Microglia 04

```{r}
ctype <- "Microglia 04"
sub_name <- "micro04"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_micro.rds")
)
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- RenameIdents(
  object = seurat, "Microglia 04" = "Microglia 05"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[10]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[10]]) <- names(seurat$subcluster_annotations)
```

# Astrocyte

```{r}
ctype <- "Astrocyte"
sub_name <- "astro"

seurat <- readRDS(
  file.path(cache_dir, "..", "step01", "processed_merged_seurat.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 2000000, dims = 1:26
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Astrocyte 01",
  "1" = "Astrocyte 02", "2" = "Astrocyte 03",
  "3" = "Astrocyte 04", "4" = "Astrocyte 05"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 26)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level1_subclusters[[3]] <- as.character(seurat$subcluster_annotations)
names(level1_subclusters[[3]]) <- names(seurat$subcluster_annotations)
```

## Astrocyte 01

```{r}
ctype <- "Astrocyte 01"
sub_name <- "astro01"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_astro.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 300000, dims = 1:29
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Astro Homeostatic 01",
  "1" = "Astro DAA 01", "2" = "Astrocyte 01",
  "3" = "Astrocyte 02", "4" = "Astrocyte 03"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 29)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[11]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[11]]) <- names(seurat$subcluster_annotations)
```

## Astrocyte 02

```{r}
ctype <- "Astrocyte 02"
sub_name <- "astro02"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_astro.rds")
)
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- RenameIdents(
  object = seurat, "Astrocyte 02" = "Astrocyte 04"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[12]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[12]]) <- names(seurat$subcluster_annotations)
```

## Astrocyte 03

```{r}
ctype <- "Astrocyte 03"
sub_name <- "astro03"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_astro.rds")
)
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- RenameIdents(
  object = seurat, "Astrocyte 03" = "Astrocyte 05"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[13]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[13]]) <- names(seurat$subcluster_annotations)
```

## Astrocyte 04

```{r}
ctype <- "Astrocyte 04"
sub_name <- "astro04"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_astro.rds")
)
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- RenameIdents(
  object = seurat, "Astrocyte 04" = "Astrocyte 06"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[14]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[14]]) <- names(seurat$subcluster_annotations)
```

## Astrocyte 05

```{r}
ctype <- "Astrocyte 05"
sub_name <- "astro05"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_astro.rds")
)
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- RenameIdents(
  object = seurat, "Astrocyte 05" = "Astrocyte 07"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[15]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[15]]) <- names(seurat$subcluster_annotations)
```

# Vascular

```{r}
ctype <- "Vascular"
sub_name <- "vasc"

seurat <- readRDS(
  file.path(cache_dir, "..", "step01", "processed_merged_seurat.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 500000, dims = 1:32
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Vasc Endothelial 01",
  "1" = "Vasc Mural 01", "2" = "Vasc Endothelial 03",
  "3" = "Vasc Endothelial 02", "4" = "Vascular 01",
  "5" = "Vasc Endothelial 04", "6" = "Vascular 02"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 32)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level1_subclusters[[4]] <- as.character(seurat$subcluster_annotations)
names(level1_subclusters[[4]]) <- names(seurat$subcluster_annotations)

level2_subclusters[[16]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[16]]) <- names(seurat$subcluster_annotations)
```

# Oligodendrocyte

```{r}
ctype <- "Oligodendrocyte"
sub_name <- "oligo"

seurat <- readRDS(
  file.path(cache_dir, "..", "step01", "processed_merged_seurat.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 2000000, dims = 1:33
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Oligodendrocyte 01",
  "1" = "Oligodendrocyte 02"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 33)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level1_subclusters[[5]] <- as.character(seurat$subcluster_annotations)
names(level1_subclusters[[5]]) <- names(seurat$subcluster_annotations)
```

## Oligodendrocyte 01

```{r}
ctype <- "Oligodendrocyte 01"
sub_name <- "oligo01"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_oligo.rds")
)
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 400000, dims = 1:35
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "Oligo Homeostatic 01",
  "1" = "Oligo DAO 01", "2" = "Oligodendrocyte 02",
  "3" = "Oligodendrocyte 01", "4" = "Oligodendrocyte 03",
  "5" = "Oligodendrocyte 05", "6" = "Oligodendrocyte 04"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

rds <- file.path(cache_dir, paste0("markers_", sub_name, ".rds"))
if (file.exists(rds)) {
  markers <- readRDS(rds)
} else {
  markers <- FindAllMarkers(seurat, only.pos = TRUE, verbose = FALSE)
  saveRDS(markers, file = rds)
}

markers <- markers[markers$p_val_adj < 0.05, ]
top <- markers %>% group_by(cluster) %>% top_n(n = -25, wt = p_val_adj)
top <- top %>% group_by(cluster) %>% top_n(n = -25, wt = avg_log2FC)
top <- merge(top, gene_anno, by = "gene")
datatable_download(top)

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 35)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[17]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[17]]) <- names(seurat$subcluster_annotations)
```

## Oligodendrocyte 02

```{r}
ctype <- "Oligodendrocyte 02"
sub_name <- "oligo02"

seurat <- readRDS(
  file.path(cache_dir, "..", "step02", "annotated_oligo.rds")
)
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- RenameIdents(
  object = seurat, "Oligodendrocyte 02" = "Oligodendrocyte 06"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level2_subclusters[[18]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[18]]) <- names(seurat$subcluster_annotations)
```

# OPC

```{r}
ctype <- "OPC"
sub_name <- "opc"

seurat <- readRDS(
  file.path(cache_dir, "..", "step01", "processed_merged_seurat.rds")
)
DefaultAssay(seurat) <- "RNA"
seurat[["SCT"]] <- NULL
gene_anno <- data.frame(
  geneid = unlist(seurat@misc), gene = rownames(seurat[["RNA"]])
)
seurat <- subset(seurat, idents = ctype)

seurat <- cluster_pipeline(
  seurat, cache_dir = cache_dir,
  sub_name = sub_name, protocol = protocol,
  vars_to_regress = vars_to_regress, parallel_override = parallel_override,
  cc = FALSE, res_divider = 10000000, dims = 1:22
)
seurat

seurat <- RenameIdents(
  object = seurat, "0" = "OPC 01"
)
seurat$subcluster_annotations <- factor(as.character(seurat@active.ident))
seurat@active.ident <- seurat$subcluster_annotations

table <- table(seurat$subcluster_annotations, seurat$allen_labels)
max <- colnames(table)[apply(table, MARGIN = 1, FUN = which.max)]
max

ElbowPlot(seurat, ndims = 50) + geom_vline(xintercept = 22)

for (i in seq_along(metadata_to_plot)) {
  if (
    "no_legend" %in% metadata_to_plot[[i]] &&
      "no_label" %in% metadata_to_plot[[i]]
  ) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_legend" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        NoLegend() +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else if ("no_label" %in% metadata_to_plot[[i]]) {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2", color = names(metadata_to_plot)[i]
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  } else {
    print(
      red_dim_plot(
        seurat, x = "umap1", y = "umap2",
        color = names(metadata_to_plot)[i], type = "cat"
      ) +
        ggtitle(names(metadata_to_plot)[i])
    )
  }
}

if (protocol[4] == "umis") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_SCT", type = "cont"
    ) + ggtitle("nFeature_SCT")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_SCT", type = "cont"
    ) + ggtitle("nCount_SCT")
  )
} else if (protocol[4] == "reads") {
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nFeature_RNA", type = "cont"
    ) + ggtitle("nFeature_RNA")
  )
  print(
    red_dim_plot(
      seurat, x = "umap1", y = "umap2", color = "nCount_RNA", type = "cont"
    ) + ggtitle("nCount_RNA")
  )
}

red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_mt", type = "cont"
) + ggtitle("percent_mt")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_ribo")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_ribo", type = "cont"
) + ggtitle("percent_malat1")
red_dim_plot(
  seurat, x = "umap1", y = "umap2", color = "percent_contam", type = "cont"
) + ggtitle("percent_contam")

VlnPlot(
  seurat,
  features = c("nFeature_RNA", "nCount_RNA"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("nFeature_SCT", "nCount_SCT"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_mt", "percent_ribo"),
  ncol = 2
)
VlnPlot(
  seurat,
  features = c("percent_malat1", "percent_contam"),
  ncol = 2
)

rds <- file.path(cache_dir, paste0("annotated_", sub_name, ".rds"))
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}

level1_subclusters[[6]] <- as.character(seurat$subcluster_annotations)
names(level1_subclusters[[6]]) <- names(seurat$subcluster_annotations)

level2_subclusters[[19]] <- as.character(seurat$subcluster_annotations)
names(level2_subclusters[[19]]) <- names(seurat$subcluster_annotations)
```

# Add Annotations

```{r}
seurat <- readRDS(
  file.path(cache_dir, "..", "step01", "processed_merged_seurat.rds")
)

seurat$level1_subclusters <- as.character(
  seurat$broad_annotations
)
names(seurat$level1_subclusters) <- names(
  seurat$broad_annotations
)
seurat$level2_subclusters <- seurat$level1_subclusters

for (i in seq_along(level1_subclusters)) {
  seurat$level1_subclusters[names(level1_subclusters[[i]])] <-
    level1_subclusters[[i]]
}
seurat$level1_subclusters <- factor(seurat$level1_subclusters)

for (i in seq_along(level2_subclusters)) {
  seurat$level2_subclusters[names(level2_subclusters[[i]])] <-
    level2_subclusters[[i]]
}
seurat$level2_subclusters <- factor(seurat$level2_subclusters)
```

```{r, fig.width = 12}
red_dim_plot(
  seurat, x = "umap1", y = "umap2", col = "level1_subclusters", type = "cat"
)
```

```{r, fig.width = 25}
red_dim_plot(
  seurat, x = "umap1", y = "umap2", col = "level2_subclusters", type = "cat"
)
```

```{r}
rds <- file.path(cache_dir, "annotated_seurat.rds")
if (!file.exists(rds)) {
  saveRDS(seurat, file = rds)
}
```

# References

This is the concluding section of the document, where we output the `sessionInfo`, and create a bibliography for works cited.

```{r}
sessionInfo()
```
